---
title: "Lacrosse Demonstration of JAGS Analysis"
author: "Roy Emanuel, Ph.D."
output: html_document  
---

# Problem

I have not yet demonstrated any of the "Bayesian" flavor of analysis,
and this is intended to be my first go at it.


The following code builds up the data used in the different models.
```{r setup, echo=FALSE}
library("tidyverse")
library("rjags")
library("readxl")
print("Hello World")
laxData <- read_excel("2015-2019 Combined Lacrosse Statistics.xlsx",
                      sheet = "Results",
                      n_max = 353)

winView <- ggplot(laxData, aes(`GA/Game`, `G/Game`))  +
    geom_point(size=10*laxData$`Win %` + 0.5)

winView2 <- ggplot(laxData, aes(`GA/Game`,
                                `G/Game`,
                                colour = `Win %`))  +
    geom_point()  +
    scale_colour_gradient(low = "yellow", high = "blue")

GoalsFor <- scale(laxData$`G/Game`)
GoalsAgainst <- scale(laxData$`GA/Game`)

GF_noAttr <- GoalsFor
attr(GF_noAttr, 'scaled:scale') <- NULL
attr(GF_noAttr, 'scaled:center') <- NULL
GA_noAttr <- GoalsAgainst
attr(GA_noAttr, 'scaled:scale') <- NULL
attr(GA_noAttr, 'scaled:center') <- NULL

laxData$GoalsFor <- as.vector(GF_noAttr)
laxData$GoalsAgainst <- as.vector(GA_noAttr)

plot(winView2)
```

The first model is a linear regression using GoalsFor and GoalsAgainst
as the independent variables and winning percentage as the dependent variable.

```{r}
mod_string <- "model {
    for (i in 1:length(y)) {
        y[i] ~ dnorm(mu[i], prec)
        mu[i] = a0 + b[1] * GoalsFor[i]  + b[2] * GoalsAgainst[i] 
    }

    a0 ~ dnorm(0.0, 1.0/1e6)
    prec ~ dgamma(5/2.0, 5*10.0/2.0)
    
    for (j in 1:2) {
        b[j] ~ dnorm(0.0, 1.0/1.0e6)
    }
    sig = sqrt( 1.0 / prec)
}"
set.seed(117)
```
The `data_jags` list uses the raw values for goals per game and goals against per game.
The `data_jags2` list uses the scaled values for goals per game.
```{r}
data_jags <- list(y=laxData$`Win %`,
                  GoalsFor=laxData$`G/Game`,
                  GoalsAgainst=laxData$`GA/Game`)

data_jags2 <- list(y=laxData$`Win %`,
                  GoalsFor=laxData$GoalsFor,
                  GoalsAgainst=laxData$GoalsAgainst)
params <- c("a0", "b", "sig")
```

The un-scaled model:
```{r}
mod <- jags.model(textConnection(mod_string), data=data_jags, n.chains=4)

update(mod, 1e3)

mod_sim <- coda.samples(model=mod,
                        variable.names=params,
                        n.iter = 5e4)

mod_csim <- as.mcmc(do.call(rbind, mod_sim))
```
The resultant diagnostic plots follow
```{r}

plot(mod_sim)
gelman.diag(mod_sim)
autocorr.diag(mod_sim)
autocorr.plot(mod_sim)
effectiveSize(mod_sim)

dic.samples(mod, n.iter=1e3)

summary(mod_sim)
```

Scaling the variables reduces the autocorrelation between draws. 

```{r}
modscaled <- jags.model(textConnection(mod_string), data=data_jags2, n.chains=4)

update(modscaled, 1e3)

modscaled_sim <- coda.samples(model=modscaled,
                        variable.names=params,
                        n.iter = 5e4)

modscaled_csim <- as.mcmc(do.call(rbind, modscaled_sim))
```
The resultant diagnostic plots follow
```{r, echo=FALSE}

plot(modscaled_sim)
gelman.diag(modscaled_sim)
autocorr.diag(modscaled_sim)
autocorr.plot(modscaled_sim)
effectiveSize(modscaled_sim)

dic.samples(modscaled, n.iter=1e3)

summary(modscaled_sim)
```
